### 2021-06-05 paper/cv_apr_2021.tex
paper files

\documentclass{article}[12pt]
\usepackage{tabularx}
\usepackage{amsfonts,amsmath}
\usepackage{verbatim} % enables you to "comment" out sections.

\emergencystretch=20pt
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% to indent
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\ind}{\hspace*{15pt}}
\newcommand{\negind}{\hspace*{-15pt}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \education{ name of school }{ dates attended }{ degrees & description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\education}[4]{\negind \begin{bf} #1 \end{bf} \hfill  #2

\hfill \parbox[b]{.93\textwidth}{ #3 }\\[-.95ex]

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \experience{ position }{ institution }{ employment period }{ description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\experience}[4]{\negind \begin{bf} #1, \end{bf} #2 \hfill  #3

\hfill \parbox[b]{.93\textwidth}{ #4 }\hfill\\[-.25ex]}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \presentation{ title }{ date }{ type & conference & location }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\presentation}[3]{\negind ``#1" \hfill  #2
\smallskip
\hfill \parbox[b]{.93\textwidth}{ #3 }
\smallskip }

\newcommand{\conference}[3]{\negind #1 \hfill  #2

\smallskip

\hfill \parbox[b]{.93\textwidth}{ #3 }

\smallskip }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\studentthesis}[3]{\negind  #1 \hfill  #2
\\
\hfill \parbox[b]{.93\textwidth}{ #3 }

\smallskip }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\student}[2]{ \begin{it} #1 \end{it} \hfill  #2

\smallskip

\hfill }

\newcommand{\publication}[6]{\parbox[b]{.93\textwidth}{\negind #1. (#2) #3, \textit{#4} #5, pages #6.}
\hfill \\[-.65ex]}

\newcommand{\book}[6]{\parbox[b]{.93\textwidth}{\negind #1. (#2) ``#3," \textit{#4} #5, #6.}
\hfill \\[-.65ex]}

\newcommand{\forthcoming}[5]{\parbox[b]{.93\textwidth}{\negind #1. (#2) ``#3," \textit{#4}}
\\[-.65ex] \hfill }

\newcommand{\inprogress}[3]{\parbox[b]{.93\textwidth}{\negind #1. ``#2" #3  }
\\[-.65ex] \hfill}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The purpose of this command is twofold:
%%   (1) Its really just a list, with no frills.
%%   (2) However, this list will justify (line up)
%%       with the formatting of the above commands.
%% Hence the name.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\justlist}[1]{\negind #1 \smallskip}

\newcommand{\courses}[3]{\negind  #1  \hfill  #2

\smallskip

\hfill \parbox[b]{.93\textwidth}{ #3 }

\smallskip }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \award{ title }{ date }{ reason }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\award}[3]{\negind #1 \hfill  #2

\smallskip

\hfill \parbox[b]{.93\textwidth}{  #3 }

\medskip }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% to resize the paper -- don't change
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addtolength{\hoffset}{-2.65cm} \addtolength{\textwidth}{5cm}
\addtolength{\voffset}{-1cm} \addtolength{\textheight}{3cm}

\pagestyle{myheadings}

\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}
\newcommand{\setlinespacing}[1]%
           {\setlength{\baselineskip}{#1 \defbaselineskip}}
\newcommand{\doublespacing}{\setlength{\baselineskip}%
                           {2.0 \defbaselineskip}}
\newcommand{\singlespacing}{\setlength{\baselineskip}{\defbaselineskip}}
\newcommand{\onehalfspacing}{\setlength{\baselineskip}%
                           {.95 \defbaselineskip}}

\begin{document}
\thispagestyle{empty}
 \onehalfspacing
%\setcounter{page}{30}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%   PERSONAL INFORMATION   %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%   (This is where you change things.)   %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
\textbf{{\large ANDREW J. LEACH, Ph.D.}}\\\smallskip \textbf{{\large Curriculum
vitae}}

\end{center}

\bigskip

{\large\scshape\hrule\vspace{0.8\baselineskip}}

\noindent  Alberta School of Business\hfill andrew.leach@ualberta.ca
\smallskip

\noindent University of Alberta \hfill http://business.ualberta.ca/aleach/
\smallskip

\noindent 3-20D Business Building \hfill Tel: 780-492-8489
\smallskip

\noindent Edmonton, Alberta T6G 2R6 (Canada)\hfill Fax: 780-492-3325
\medskip

{\large\scshape\hrule\vspace{0.8\baselineskip}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
                                                           %%
\medskip
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%   Current Position %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large AFFILIATIONS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \experience{ position }{ institution }{ employment period }{ description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\experience{Alberta School of Business}{University of Alberta}{2011--}{Associate Professor, with tenure} %\vspace{-2\baselineskip}

\experience{Smart Prosperity}{Research Network}{2018--}{Member}

\experience{CABREE}{Centre for Applied Business Research in Energy and the Environment}{2014--}{Academic Director} %\vspace{-2\baselineskip}

\experience{CIRANO}{Center for Interuniversity Research and Analysis on Organizations}{2005--}{Researcher} %\vspace{-2\baselineskip}

%\experience{CIRPE\'{E}}{Centre interuniversitaire sur le risque, les politiques \'{e}conomique et l'emploi}{March, 2005-present}{Researcher} %\vspace{-2\baselineskip}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

                                                           %%
\bigskip                                                            %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%   Current Position %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large PAST AFFILIATIONS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\bigskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \experience{ position }{ institution }{ employment period }{ description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\experience{Government of Alberta}{Department of Environment and Climate Change}{2015}{Chair, Climate Leadership Panel} %\vspace{-2\baselineskip}

\experience{Alberta School of Business}{University of Alberta}{2013--2015}{Enbridge Professor of Energy Policy} %\vspace{-2\baselineskip}

\experience{Environment Canada}{Strategic Policy Branch}{2012--2013}{Visiting Scholar} %\vspace{-2\baselineskip}

\experience{Sustainable Prosperity}{Research Network}{2008--2017}{Member}
%\vspace{-2\baselineskip}

\experience{Alberta School of Business}{University of Alberta}{2006-- 2011}{Assistant Professor} %\vspace{-2\baselineskip}

\experience{HEC Montr\'{e}al}{Institute of Applied Economics
}{2003--2007}{Assistant Professor} %\vspace{-2\baselineskip}

\experience{Queen's University}{Economics Department}{2002 -- 2003} {Teaching
Fellow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

                                                           %%
\bigskip                                                            %%

%% This says the header should be my name (and the page number)    %%%
\markright{ANDREW J. LEACH} \vspace{1cm}                                 %%%

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%   EDUCATION   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large EDUCATION}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\bigskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \education{ name of school }{ dates attended }{ degrees & description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\education{University of Alberta}{2021} {Master of Laws}

\education{Queen's University}{2004} {Ph.D.(Economics)}

 \education{University of Guelph}{1999} {Master of Arts (Economics)}

 \education{University
of Guelph}{1998} {Bachelor of Science (Environmental Science)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% I like to start a new page around here, but you can move this. %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%   PROJECTS   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\noindent {\bf \large RESEARCH PROJECTS} $_{\begin{tabular}{l}      %%
%\hline \hspace{351pt} \\                                            %%
%\end{tabular}}$                                                     %%
                                                                    %%
%\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \presentation{ title }{ date }{ additional information }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\presentation{Single-Variable Linear Equations with Integer
%Coefficients}{September 20xx -- Spring 20xx}
%{ Senior Thesis, Department of Mathematics, Undergraduate Institution. \\
%Advisor: Professor Matt Amatician }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\medskip                                                            %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%   PUBLICATIONS   %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent \textbf{{\large PUBLICATIONS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\inprogress{Nigel Bankes, Andrew Leach and Martin Olszynski (forthcoming)}{Case Comment: \textit{References re Greenhouse Gas Pollution Pricing Act}, 2021 SCC 11}{\textit{Energy Regulation Quarterly}}

\publication{Leach, Andrew}{2021}{Environmental Policy is Economic Policy:
Climate Change Policy and the General Trade and Commerce Power}{\textit{Ottawa Law Review}}{52(2)}{97-156}

\publication{Leach, Andrew and Nic Rivers and Blake Shaffer}{2020}{Canadian Electricity Markets during the COVID-19 Pandemic: An Initial Assessment}{\textit{Canadian Public Policy}}{46(S2)}{S145-S159}

\publication{Olszynski, Martin, Nigel Bankes, and Andrew Leach}{2020}{Case Comment: \textit{Reference re Greenhouse Gas Pollution Pricing Act}, 2020 ABCA 74}{\textit{Journal of Environmental Law and Practice}}{33(2)}{159-184}

\publication{Leach, Andrew and Eric M. Adams}{2020}{Seeing Double: Peace, Order, and Good Government, and the Impact of Federal Greenhouse Emissions Legislation on Provincial Jurisdiction}{\textit{Constitutional Forum}}{29 (1)}{1-20}

\publication{Bo\v skovi\'{c}, Branko and Andrew Leach}{2020}{Leave it in the ground? Oil sands development under carbon pricing}{\textit{Canadian Journal of Economics}}{53(2)}{526-562}

\publication{Heyes, Anthony, Andrew Leach and Charles F. Mason}{2018}{The Economics of the Canadian Oil Sands}{\textit{Review of Environmental Economics and Policy}}{12 (2)}{242-263}

\publication{Leach, Andrew}{2016}{Canada in a Low-Carbon World: Impacts on new and existing resources}{\textit{Canadian Public Policy}}{42 (S1)}{S18-S23}

\publication{Leach, Andrew}{2015}{Canadian Hydrocarbon Resources in an Era of Manufactured Energy”, Chapter in Roland Paris and Taylor Owen (eds) The World Won't Wait: What Canada needs to rethink its international policies}{University of Toronto Press}{2015}{240}

\publication{Leach, Andrew and Branko Boskovic}{2014}{Carbon cost will not stop oil-sands work}{\textit{Nature} Letters}{511}{page 534-534}

\publication{Leach, Andrew}{2012}{Alberta's Specified Gas Emitters
Regulation}{\textit{Canadian Tax Journal}}{60(4)}{881-898}

\publication{Chakravorty, Ujjayant, Andrew Leach, and Michel
Moreaux}{2012}{Cycles in nonrenewable resource prices with pollution and
learning-by-doing}{\textit{Journal of Economic Dynamics and
Control}}{36(10)}{1448-1461}

\publication{Laurent-Lucchetti, Jeremy and Andrew Leach}{2011}{Generational
welfare and induced innovation under climate change
policy}{\textit{Scandinavian Journal of Economics}}{113(4)}{904-936}

\publication{Leach, Andrew, Charles F. Mason and Klaas van't
Veld}{2011}{Co-optimization of oil production and carbon
sequestration}{\textit{Resource and Energy Economics}}{33}{893-912}

\publication{Leach, Andrew, Joseph Doucet and Trevor Nickel}{2011}{The
effectiveness of Alberta biofuels policy}{\textit{Energy Policy}}{39(7)}
{4007-4015}

%\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}
%TENURE
%\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}

\publication{Chakravorty, Ujjayant, Andrew Leach, and Michel
Moreaux}{2011}{Would Hotelling Kill the Electric Car?}{\textit{Journal of
Environmental Economics and Management}}{61(3)}{281-296}

\book{Chakravorty, Ujjayant, Shelby Gerking and Andrew Leach}{2010}{State Tax
Policy and Oil Production: The Role of the Severance Tax and Credits for
Drilling Expenses}{US Energy Tax Policy,}{Metcalf, Gilbert (ed), Cambridge
University Press}{414 pages.}

\publication{Leach, Andrew}{2009}{The Welfare Implications of Climate Change
Policy}{Journal of Environmental Economics and Management}{57(2)}{151-165.}

\publication{Bahn, Olivier and Andrew Leach}{2008}{The Secondary Benefits of
Climate Change Mitigation: An overlapping generations approach}{Computational
Management Science}{5(3)}{233-257}

\publication{Leach, Andrew}{2007}{The Climate Change Learning Curve}{Journal of
Economic Dynamics and Control}{33}{1728-1752}

\publication{Clark, C. Robert and Andrew Leach}{2007}{The Potential for
Electricity Market Restructuring in Quebec}{Canadian Public Policy}
{33(1)}{1-20}

\publication{Leach, Andrew}{2006}{Regulation by Reputation}{Computers and
Operations Research}{33}{409-425}

\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%   PROJECTS   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large CURRENT PROJECTS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\inprogress{Andrew Leach}{“Economic Policy is Environmental Policy”}{Accepted for publication in the \textit{Ottawa Law Review}}

\inprogress{Andrew Leach}{The No More Pipelines Act? Bill C-69’s impact on oil sands pipelines}{Submitted, \textit{Alberta Law Review}}

\inprogress{Andrew Leach}{The Constitutional Limits on Federal GHG Emissions Policies}{In preparation for submission to the \textit{Canadian Journal of Economics}}

\inprogress{Fenner L Stewart, Andrew Leach and Jennifer Winter}{A Federal Backstop to Provincial Carbon Pricing: Decarbonization in an Age of Uncooperative Federalism}{Manuscript}

\inprogress{Andrew Leach and Blake Shaffer}{Carbon Price Pass-Through in Alberta's Electricity Market}{In preparation for presentations}

\inprogress{Branko Boskovic, Andrew Leach and Charles F. Mason}{Nonrenewable resources, reserves, and carbon pricing}{In preparation for submission}

\inprogress{Andrew Leach, and Alain-Desir\'{e} Nimubona}{Technology search and environmental regulation.}{In preparation for re-submission}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%   PROJECTS   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent \textbf{{\large SELECTED PROFESSIONAL AND ADVISORY ACTIVITY}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\smallskip %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\inprogress{Leach, Andrew and Angela Adams, Stephanie Cairns, Linda Coady and Gordon Lambert}{Climate Leadership Panel, Report to the Minister}{Prepared for Alberta Environment and Climate Change, 2015.}

\inprogress{Leach, Andrew}{Royalty Review Scoping Document}{Prepared for Alberta Energy, 2015.}

\inprogress{Leach, Andrew}{Emerging oilsands technology and prospects for GHG
	emissions reductions}{Prepared for Environment Canada, 2012.}

\inprogress{Leach, Andrew}{An evaluation of `A Climate Change Plan for the
Purposes of the Kyoto Protocol Implementation Act (Bill C-288)'}{prepared for
the National Roundtable on the Environment and the Economy annually from
2007-2011.}

\inprogress{Leach, Andrew}{GHG Emissions Policies for Alberta: Current
strengths and weaknesses, key future threats, and a proposed policy solution
for the Province of Alberta}{Prepared for Alberta Environment, July, 2009.}

\inprogress{Leach, Andrew and Dave Sawyer}{An evaluation of the modeling and
information transmittal process used for the development of ``Alberta's 2008
Climate Change Strategy"'}{Prepared for the Office of the Auditor General of
Alberta, July, 2008.}

\inprogress{Leach, Andrew and C. Robert Clark}{The determinants of demand for
natural gas: a review of the literature}{Prepared for Gaz M\'{e}tropolitain,
July, 2006.}

\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Research and Teaching Grants %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large MAJOR GRANTS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\justlist{\begin{bf}Canada First Excellence Research Fund\end{bf}
	Future of Energy Systems Institute, University of Alberta, \$75,000,000 over 7 years, 2016.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Insight Development Grant 430-2014-0275, \$58,630 over 2 years, 2014.}

\justlist{\begin{bf}Sustainable Prosperity\end{bf} Project grant with Branko
Boskovic, \$8000, 2014.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Insight Development Grant 430-2012-0079, \$45,612 over 2 years, 2012.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Collaborator, Alberta Climate Dialogue (ABCD), Community-University Research
Alliances,\$1 million over 6 years, 2010.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Standard Research Grant, Grant Number 410-2009-1044, \$18600 over 3 years,
2009.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Sustainable Prosperity Network Grant, \$1.8 million over 7 years, 2008.}

\justlist{\begin{bf}Fonds qu\'{e}becois de la recherche sur la soci\'{e}t\'{e}
et la culture (FQRSC)\end{bf} Team Grant, \$332 200 over 4 years, with G.
Dionne, R. Gagn\'{e}, P. Lanoie, P. Fran\c{c}ois, and B. Dostie, 2005.}

%During the course of the year, I have given a number of academic presentations %and invited lectures as follows:
%
%University of Ottawa, Economics of Environmental Policy Class Presentation, %July 23
%University of Ottawa, Economics of Environmental Policy Workshop, July 25
%Presentation to Ivey School of Business Students, UWO, October 3
%AIMCO Investor Conference, October 16
%Better Policy for a Better World: Why good ideas aren't enough, Max Bell School %of Public Policy, McGill, November 14
%Federal Cabinet Retreat 2020, Winnipeg, January 20
%Rural Economy and Environmental Sociology (REES) department seminar, February 7
%Canadian Bar Association, Edmonton Environmental Law Chapter, February 13
%Alberta Corporate Growth Summit, March 11
%Harry Ainlay High School, March 13
%UBC Political Science class, April 1
%Canadian Bar Association, Edmonton Constitutional Law Chapter, May 7
%Alberta Electrical System Operator (AESO) University, June 5.
%Economic Society of Northern Alberta, June 16

%Writing

%Alberta Court of Appeal Opines That Federal Carbon Pricing Legislation %Unconstitutional (with Martin Olszynski and Nigel Bankes), ABlawg, March 17, %2020

%Clearing the Air on Teck Frontier (Extended ABlawg Edition) (with Martin %Olszynski), ABlawg, February 13, 2020

%Clearing the air on Teck Frontier: Here are the expected benefits — and harms — %of the oilsands project (with Martin Olszynski), CBC, Feb 12, 2020

%Alberta's TIER regulations good on electricity, not so good on oilsands
%CBC, October 30, 2019

%Election 2019: A national reckoning on climate change
%CBC, October 16, 2019

%How The Four Federal Parties' Climate Plans Stack Up (with Katharine Hayhoe), %Chatelaine, October 3, 2019

%Liberals' climate balancing act: The world is watching, CBC September 25, 2019

%NDP climate plan tries (and fails) to carve out middle ground, CBC,September %13, %2019

%What's really holding back the oilsands? It's not the bill of goods you're being %sold, CBC, September 4, 2019

%Conservative climate plan is cloaked in mystery, choked with irony, CBC,
%August 12, 2019

%Greens' climate plan adds up to Mission: Improbable, CBC, August 12, 2019

\end{document}