### 2022-01-18 electricity_deck.Rmd
updating

---
title: Alberta's Electricity Market
subtitle: Energy Industries and Markets Slide Pack
author: 
  - Andrew Leach 
institution: 
- Alberta School of Business and CABREE
fontsize: 10pt
output:
 beamer_presentation:
   #template: "C:/Users/aleach/Google Drive/default_beamer.tex"
   #Mac
   template: "ajl-latex-beamer.tex"
    #PC
   #template: "C:/Users/aleach/Google Drive/ajl-latex-beamer.tex"

   theme: "Darmstadt"  
   incremental: false
   slide_level: 2
    #keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)

library(ggplot2)
library(lubridate)
library(dplyr)
library(viridis)
library(zoo)
```

#Introduction

## Why should electricity interest you?

Price volatility:

- Electricity prices are an order of magnitude more volatile than oil or gas prices

- Both supply and demand in Alberta affect us more than global or regional supplies and demands in oil and gas markets

New technology:

- Electricity is, arguably, changing faster than any other energy market

- Alberta's electricity market is entering a period of market- and regulatory-driven transition

Economics 101 in action

- Nowhere else will you see supply and demand curves actually mapped out in real time determining prices as clearly as in Alberta's power market

```{r plotit, echo=FALSE,results=FALSE,message=FALSE}
options(warn=-1) #suppress warnings in Markdown

slide_theme<-function(){
  return( theme(panel.border = element_blank(),
        panel.grid = element_blank(),
        panel.grid.major.y = element_line(color = "gray",linetype="dotted"),
        axis.line.x = element_line(color = "gray"),
        axis.line.y = element_line(color = "gray"),
        axis.text = element_text(size = 20),
        axis.text.x = element_text(margin = margin(t = 10)),
        axis.title = element_text(size = 20),
        #axis.label.x = element_text(size=20,vjust=+5),
        plot.subtitle = element_text(size = 20,hjust=0.5),
        plot.caption = element_text(face="italic",size = 16,hjust=0),
        legend.key.width=unit(2,"line"),
        legend.position = "bottom",
        #legend.direction = "horizontal",
        #legend.box = "horizontal",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        plot.title = element_text(hjust=0.5,size = 24)))
}

```

```{r gendata, echo=FALSE,results=FALSE,message=FALSE}

load("gen.RData") ## which is here *equivalent* to

gen_data<-gen_data[-is.na(gen_data$gen),]

gen_data$AIL<-as.numeric(gen_data$AIL)
gen_data$Price<-as.numeric(gen_data$Price)
gen_data$Demand<-as.numeric(gen_data$Demand)

```

```{r, echo=FALSE,warnings=FALSE,results=FALSE,message=FALSE}
filename<-paste("forecast_data",".RData",sep = "")
#save(data_store, file= filename) 
load(file = filename)

load("all_merit.RData")  #load merit order data

#hourly prices  
forecast_data$day_ahead_forecasted_ail<-gsub(",","",forecast_data$day_ahead_forecasted_ail)
forecast_data$actual_ail<-gsub(",","",forecast_data$actual_ail)
forecast_data$forecasted_actual_ail_difference<-gsub(",","",forecast_data$forecasted_actual_ail_difference)
forecast_data$Year<-year(forecast_data$date)
forecast_data$actual_posted_pool_price<-as.numeric(forecast_data$actual_posted_pool_price)
forecast_data$forecast_pool_price<-as.numeric(forecast_data$forecast_pool_price)
forecast_data$actual_ail<-as.numeric(forecast_data$actual_ail)
forecast_data$day_ahead_forecasted_ail<-as.numeric(forecast_data$day_ahead_forecasted_ail)
sub_samp<-forecast_data
sub_samp$time<-NULL
sub_samp<-sub_samp[!is.na(sub_samp$actual_posted_pool_price),]
df1 <- sub_samp %>% group_by(he,Year) %>% summarise(AIL=mean(actual_ail),p_mean=mean(actual_posted_pool_price))

#png(file="hourly-prices.png", width = 1400, height = 750)
```

## Price volatility
```{r, fig.width=16,fig.height=11, echo=FALSE,warnings=FALSE,results=FALSE,message=FALSE}
p<-ggplot(df1, aes(he,p_mean,group=as.factor(Year),colour=as.factor(Year))) +
  geom_line(data=df1,size=1.5) +
  geom_line(data=subset(df1,Year==2017),size=3) +
  #geom_point(size=1.5) +
  labs(colour="Year") +
  scale_color_viridis(discrete=TRUE)+   
  theme_classic() +
  theme(panel.border = element_blank(),
        panel.grid = element_blank(),
        panel.grid.major.y = element_line(color = "gray",linetype="dotted"),
        axis.line.x = element_line(color = "gray"),
        axis.line.y = element_line(color = "gray"),
        axis.text = element_text(size = 20),
        axis.text.x = element_text(margin = margin(t = 10)),
        axis.title = element_text(size = 20),
        #axis.label.x = element_text(size=20,vjust=+5),
        plot.subtitle = element_text(size = 20,hjust=0.5),
        plot.caption = element_text(face="italic",size = 16,hjust=0),
        legend.key.width=unit(2,"line"),
        legend.position = "bottom",
        #legend.direction = "horizontal",
        #legend.box = "horizontal",
        legend.text = element_text(size = 20),
        plot.title = element_text(hjust=0.5,size = 24))+
  labs(y="Hourly Average Power Price ($/MWh)",x="Hour",
             title="Hourly Average Energy Prices ($/MWh, 2000-2018)",
             caption="Source: AESO Data, Accessed via NRGStream, Graph by Andrew Leach")

print(p)
```

## Price volatility
```{r, fig.width=16,fig.height=11, echo=FALSE,warnings=FALSE,results=FALSE,message=FALSE}
sub_samp$month<-month(sub_samp$date)
sub_samp$Date_ID = as.yearmon(sub_samp$date)
df1 <- sub_samp %>% group_by(Date_ID) %>% summarise(cons =sum(actual_ail)/1000,AIL =mean(actual_ail),Peak_AIL =max(actual_ail), p_mean=sum(actual_posted_pool_price*actual_ail)/sum(actual_ail))
df1$Date<-as.Date(df1$Date_ID)
df1$m12_avg<-as.numeric(rollapply(df1$p_mean,12,mean,fill=NA,align = c("right")))
df1$m12_ail_avg<-as.numeric(rollapply(df1$AIL,12,mean,fill=NA,align = c("right")))

lims=c(min(df1$Date)+years(1),max(df1$Date)+months(3))
p<-ggplot(data=subset(df1,Date>="2001-01-01"), aes(Date,p_mean,colour="A"),size=2.5) +
  geom_line(size=1.5) +
  geom_line(data=subset(df1,Date>="2001-01-01"),aes(Date,m12_avg,colour="B"),size=2.5) +
  scale_color_viridis("",labels=c("Monthly Average Prices","12 Month Rolling Average Prices"),discrete=TRUE)+   
  scale_x_date(name=NULL,date_breaks = "1 year", date_labels =  "%b\n%Y",limits=lims,expand=c(0,0)) +
  theme_classic() +
  slide_theme()+
      labs(y="Monthly/Annual Average Power Price ($/MWh)",x="\nMonth",
             title="Load-Weighted Monthly Average Energy Prices ($/MWh, 2005-2017)",
             caption="Source: AESO Data, Accessed via NRGStream, Graph by Andrew Leach")

print(p)
```

## Supply and Demand

```{r, fig.width=16,fig.height=11, echo=FALSE,warnings=FALSE,results=FALSE,message=FALSE}

df1<-subset(merit_data, he==18 & date==as.Date("2017-10-05"))

p<-ggplot(df1,aes(merit,price,fill=offer_sum)) +
  geom_rect(mapping=aes(xmin=merit-size,xmax=merit,ymin=-20,ymax=price))+
  scale_fill_viridis("Offer\nControl",discrete = TRUE)+   
  theme_classic() +
  slide_theme()+
  labs(x=paste("Offered Generation (MW)"),y="Price ($/MWh)",
       title=paste("Alberta Energy Merit Order, ",max(df1$date)," ",max(df1$he),":00",sep = ""),
       caption="Source: AESO Data, graph by Andrew Leach.")
print(p)

```

## New Technology

```{r,echo=FALSE,message=FALSE, out.width = "320px"}
knitr::include_graphics("/Users/aleach/Google Drive/eclipse.png")

```

#The Market

## The Grid 2.0
```{r,echo=FALSE,message=FALSE, out.width = "320px"}
knitr::include_graphics("grid_ess.png")
```

## Market Participants
\begin{itemize}
\setlength\itemsep{2em}
\item Generation

\item Transmission

\item Distribution

\item Ancillary Services

\item Load

\item Storage
\end{itemize}

## Energy units - electricity 
Watts: measure of capacity (instantaneous production, installed capacity, or instantaneous demand)

- Alberta system demand: 7,200-10,700 MW (million watts)

- Capital Power's Genessee 3 power plant is: 450 MW

Watt hours: measure of energy (production or demand during a given period of time; i.e. flow through)
    -Production over a day, week, month, year

Volts are a measure of the electrical potential - the ability to convert charge to power: Watts=amps x volts
    -Transmission lines: 150-765 kV
    -Distribution lines: 13,800 Volts
    -Household wiring: 120-240 Volts

# Prices
## Energy Prices
Electricity prices are expressed in power delivered over time
Cents/kilowatt-hour (c/kWh)
Dollars per megawatt-hour ($/MWh)
Capacity costs are expressed in a cost per megawatt or cost of capacity
Genessee 3 cost approximately $1.5 million/MW or $1.50 per watt to build
Solar panel prices have declined to now lie under $1/W of capacity
Balance of system costs imply that a solar system costs $2-3/W of installed capacity
Levelized costs of electricity (supply costs) in $/MWh

## Other Prices
Other prices matter for electricity markets as well
Renewable energy credits (usually prices in $/MWh)
Emissions credits or permits ($/tonne)
Capacity payments ($/MW)
Air emissions permits or credits ($/tonne)

#Market Participants

## Test

#Generation Sources

## Test

#GHG Emissions

## Test

#Etc

## Test